---
title: "Exploring Climate Impact in Saudi Arabia  "
output: html_document
---

```{r setup, include=FALSE}
# هذا تشنك عادي اسمه setup
library(ggplot2)
library(dplyr)
```

```{r open_site, echo=FALSE, include=FALSE}
browseURL("https://raghadaltassan.github.io/climatechange/")
```
### Problem of the Study
Climate change has emerged as a global environmental concern, with rising concentrations of greenhouse gases (GHGs)—notably carbon dioxide (CO₂) and methane (CH₄)—accelerating global warming. Saudi Arabia, being one of the world’s top oil producers, faces unique climate challenges due to its arid conditions and industrial activities. This study investigates how CO₂ and CH₄ emissions contribute to temperature changes across the Kingdom using satellite-derived data and spatial analysis.

---

### Why It Matters
Understanding the relationship between greenhouse gas emissions and temperature trends is essential for tackling climate-related risks. Saudi Arabia is especially vulnerable due to its extreme climate, rapid urban growth, and heavy reliance on fossil fuels. This research provides insights that support national environmental policies and sustainable development goals—especially those set by the Saudi Green Initiative and Vision 2030. It also empowers decision-makers with scientific evidence to reduce emissions and adapt to climate change effectively.

---

### Climate Trends in Saudi Arabia
The Kingdom is characterized by harsh desert conditions, low vegetation cover, and high industrial emissions. These factors make it a critical region for understanding greenhouse gas dynamics and their influence on rising surface temperatures. By analyzing temporal and spatial variations from 2018 to 2021, this dashboard captures emission patterns and temperature shifts at the city level.

---

### Greenhouse Gas Distribution
- **CO₂ Emissions**: Primarily result from fossil fuel combustion and industrial activity.  
- **CH₄ Emissions**: Linked to agriculture, landfills, and natural gas infrastructure.  
- **Temperature Change**: Increasing surface temperatures indicate intensified heatwaves and environmental vulnerability.

---

### Research Focus and Objectives

This study focuses on understanding the impact of greenhouse gas (GHG) emissions—specifically carbon dioxide (CO₂) and methane (CH₄)—on climate variability in Saudi Arabia. Given the Kingdom’s arid climate, urban expansion, and heavy reliance on fossil fuels, the research investigates how rising GHG concentrations influence regional temperature patterns.

The key objectives of this research are to:

1. Analyze the **annual spatial and temporal variability** of CO₂ and CH₄ concentrations across Saudi Arabian cities from 2018 to 2021.  
2. Examine the **relationship between greenhouse gas emissions and climate change indicators**, particularly land surface temperature.  
3. Utilize **satellite-derived data and advanced spatial modeling techniques** (e.g., Bayesian Hierarchical Models with INLA) to provide high-resolution assessments of GHG dynamics.  
4. Offer **scientific insights that support national and international climate policies**, such as Saudi Arabia’s Vision 2030 and the Saudi Green Initiative.
This project focuses on Saudi Arabia—a region with high emissions and environmental vulnerability. The key research objectives are:  
1. Analyze the spatial and temporal patterns of carbon dioxide (CO₂) and methane (CH₄) emissions across Saudi Arabia from 2018 to 2021.  
2. Examine how GHG concentrations relate to variations in surface temperature.  
3. Visualize emission patterns and temperature changes using interactive maps.  
4. Identify regional disparities in environmental risk.  
5. Provide insights that support national policies for emissions reduction and climate adaptation.

---

### Research Significance
By understanding where and how greenhouse gases accumulate—and how they affect temperature—this research helps prioritize regions for environmental action. The interactive dashboard delivers data-driven insights that are crucial for evidence-based climate policy, regional planning, and national sustainability strategies. The findings also contribute to Saudi Arabia’s climate commitments under Vision 2030.
